---
title: "subsetting all the shit"
author: "a dumbass"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    highlight: tango
    toc: 3
    toc_float: yes
    number_sections: true
    code_folding: show
    code_download: true
subtitle: "Reproducible Report"
link-citations: yes
fontsize: 12pt
bibliography: references.bib
csl: plos-genetics.csl
---

```{r packages, echo=FALSE, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = './')
library(BiocManager)
library(devtools)
#install.packages("devtools")
#install.packages("devtools")

#BiocManager::install("dada2")
library(dada2)
#BiocManager::install("DECIPHER", force=TRUE)
library(DECIPHER)
#install.packages("phangorn")
library(phangorn)
#BiocManager::install("phyloseq", force=TRUE)
library(phyloseq)
#BiocManager::install("decontam")
library(decontam)
#install.packages("tidyverse")
library(tidyverse)
library(dplyr)
#BiocManager::install("multtest", force=TRUE)
#devtools::install_github("gauravsk/ranacapa")
library(ranacapa)
#BiocManager::install("genefilter")
library(genefilter)
#install.packages("BiodiversityR")
library(BiodiversityR)
library(vegan)
library(reshape2)
#devtools::install_github("jbisanz/qiime2R")
library(qiime2R)
#install.packages("dplyr")
library(dplyr)
#install.packages("parallelDist")
library(parallelDist)

#Install the following to use National Parks color palette for plots
#install.packages("devtools")
#devtools::install_github("katiejolly/nationalparkcolors")
library(NatParksPalettes)

#Install the following to use Knit to PDF option:
#install.packages(tinytex)
#tinytex::install_tinytex()

library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(kableExtra)

#Install the following to use Knit to PDF option:
#install.packages(tinytex)
#tinytex::install_tinytex()

library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(lmerTest)
cite(lmerTest)

## Cite all packages and save into a .bib file titled "packages.bib".
knitr::write_bib(.packages(), file = "packages.bib")
```

```{r setup, include = FALSE, cache = FALSE, message = FALSE}

# Require knitr for setting up global options
require('knitr')

# Set global options
knitr::opts_chunk$set(echo = TRUE)

# Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60),
    highlight = TRUE)

# Caching code
opts_chunk$set(cache = 2, cache.path = "cache/")

# Saves plots in png and pdf format at 300 dpi in folder "generated_plots"
opts_chunk$set(fig.path = "./generated_plots/", dev = c("png", "pdf"),
    dpi = 300)

# Place figures as close as possible to code in which it is called
opts_chunk$set(fig.pos = "H")

```

# Downstream Analysis in R

## Global Options
**Be sure to set up your project this workflow by establishing a working directory and global code chunk options in your second code chunk:*

## Input QIIME2 Files
QIIME2 artifacts were uploaded into R RStudio 4.2.2 for simplified downstream filtering and analyses.

Load non-rarefied QIIME2 metadata file: "SageGrouse_metadata.csv"

```{r import_files_r}
#This is extracting embedded data and object metadata into an R session, in other words, asking R to pull only the data from the list and put it in an asv table, displaying species and quantity
SVs <- qiime2R::read_qza("./table.qza")
#
#"$" allows you to access a single variable in a data set
asv_table<-(SVs$data)
#
# Save data frame into "Asv_table
Asv_table<-data.frame(asv_table)
#
#change column name in new table
colnames(Asv_table) <- colnames(asv_table) 
#
#Reading in evolutionary tree, breaking down taxonomic ranks
tree<- qiime2R::read_qza("./sepp-tree.qza") 
#
#Save tree data into vector named "Tree"
Tree<-tree$data#
##Reading in Metadata
Metadata_final <- read.csv("./Metadata.csv")
#
#Copy only the sample names as a character list, and set them as the row labels for the metadata sheet, identifying that "sample_name" is not a number
metadata_finalnames <- as.character(Metadata_final$sample_name)
row.names(Metadata_final) <- Metadata_final$sample_name
#
```

```{r metadata}
# make table with knitr and KableExtra, use "scroll_box" option to add a scroll bar due to the size of this table
knitr::kable(Metadata_final, caption = "Table 1. Metadata") %>%
    kable_styling(c("striped", "scale_down", "condensed")) %>% scroll_box(width = "750px", height = "500px") 

```

Next, save taxonomy .qza's into vectors as well.
```{r Qiime2_taxonomy}
# Read taxonomy .qza into vector
taxonomy<-read_qza("./taxonomy.qza") 
#
#When taxonomy is imported, a single string is returned along with a confidence score. For many analyses, we will want to break up this string, and for that purpose, the parse_taxonomy() function is provided
taxonomy<-qiime2R::parse_taxonomy(taxonomy$data)
#
#Again, convert names to characters, so they will now be read as characters rather than data values
taxonomy$Kingdom <- as.character(taxonomy$Kingdom) 
taxonomy$Phylum <- as.character(taxonomy$Phylum)
taxonomy$Class <- as.character(taxonomy$Class)
taxonomy$Order <- as.character(taxonomy$Order)
taxonomy$Family <- as.character(taxonomy$Family)
taxonomy$Genus <- as.character(taxonomy$Genus)
taxonomy$Species <- as.character(taxonomy$Species)
taxonomy <- as.matrix(taxonomy)
```

## Create Phyloseq Object {#phyloseq}

Create a phyloseq object using all of the files we got from QIIME2. A **phyloseq object** is a data format commonly used for microbiome data in R.

```{r phyloseq_object}
#Note that this object is created with the non-rarefied data (uses the ASV table with unique sequences we pulled from out dada2 output at start of second chunk)
phyloseq_16S <- phyloseq::phyloseq(otu_table(Asv_table, taxa_are_rows = TRUE, 
errorIfNULL = TRUE), sample_data(Metadata_final), tax_table(taxonomy), 
phy_tree(Tree))
#
#Check to make sure your phyloseq object was properly loaded, if so, the run will summarize the phyloseq object in your console, showing the number of ASVs and samples contained in the object. Note that it is useful to periodically summarize the object to ensure that all desired information is still available
#phyloseq_16S  
#
#saveRDS is a function to write a single R object to a file. To restore the file back into an object, use command "readRDS()"
saveRDS(phyloseq_16S, "PH_phyloseq_16S.rds")
#
#Takes sample data from phyloseq_16S and created a file called "metadata", available in the top right quadrant called "Data Environment", so now we can select it and view all of the metadata in a chart
metadata <- data.frame(sample_data(phyloseq_16S)) 
#
##ASV table containing abundances of each ASV
#The column headers contain full ASV sequences, row names indicate sample names, and observations show ASV counts
ASV_table <- data.frame(otu_table(phyloseq_16S)) 
#
#changed column names from asv_table to ASV_table
colnames(ASV_table) <- colnames(Asv_table) 
#
##Taxonomic table containing taxonomy for each ASV sequence
#Row names contain full ASV sequences whereas columns show taxonomic identification at different resolutions
#Most, but not all ASVs will be identified down to genus level
taxa_table <- data.frame(tax_table(phyloseq_16S))
#
knitr::kable(taxa_table, caption = "Table 2. Initial Taxonomy") %>%
    kable_styling(c("striped", "scale_down", "condensed")) %>% scroll_box(width = "750px", height = "500px") 
```

### Decontaminating Data & Taxonomy Subset {#decontam}

We are sub-setting here for bacteria only, and filtering out mtDNA and chloroplast DNA.

**Note: We also took the opportunity to remove UNK/Mansfield locations from here, for now due to low sample counts.**

```{r decontam}
#Running this gives a summary look at how many samples, OTU, taxa etc. just checking to see how things have changed since the last run
phyloseq_16S 
#
#Remove archea and unassigned
phyloseq_16s_bact<-subset_taxa(phyloseq_16S, Kingdom == "d__Bacteria")
#Summarize to verify that archea was removed and only bacteria remains, so expect to see a decline in taxa groups identified
phyloseq_16s_bact 
#
#Removing the mitochondria and chloroplasts, "!" means to exclude or remove
phyloseq_16s_bact2 <-subset_taxa(phyloseq_16s_bact, Family!= "Mitochondria" | 
is.na(Family) & Class!="Chloroplast" | is.na(Class) )
#
#summarize.

phyloseq_16s_bact2

```
### Rarefy data {#rarefy}

Here, we select a threshold to rarefy to. **Rarefaction** is the process of selecting a specific number of samples that is equal or less than the number of samples in the smallest sample, and then randomly discarding reads from larger samples until the number of remaining samples is equal to the threshold. 

```{r rarefy, echo = T, results = 'hide'}
#Look at how many reads are in each sample
sort(colSums(otu_table(phyloseq_16s_bact2)))
#Remove all samples that did not meet 2500 threshold
phyloseq_16S_rarifed <- subset_samples(phyloseq_16s_bact2, sample_name != "21-R12")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "23-F26")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "22-R15 ")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "22-R12")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "23-F121")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4D")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "22-R13")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "23-F72")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "23-F66")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4G")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4C")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank5.5C")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "23-F73")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank5.5G")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "23-F124")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4E")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank5.5H")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4F")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank5.5E")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank5.5D")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank5.5F")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4H ")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank6")
#
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank7")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank8")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank4")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank1")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank2")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank3")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "Blank5")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "22-R15")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "12578.blank4.4H")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "PCR_Pos1")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "PCR_Pos2")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "PCR_Neg1")
phyloseq_16S_rarifed <- subset_samples(phyloseq_16S_rarifed, sample_name != "PCR_Neg2")
#Summarize to verify that samples were removed, as well as sampling depth = 2500
sort(colSums(otu_table(phyloseq_16S_rarifed)))
#
```
```{r rarefy, echo = T, results = 'hide'}
#Look at how many reads are in each sample
sort(colSums(otu_table(phyloseq_16s_bact2)))

Metadata_reduced <- read.csv("./Metadata.csv")
#Remove all samples that did not meet 2500 threshold
Metadata_reduced <- subset(Metadata_reduced, sample_name != "21-R12")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "23-F26")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "22-R15 ")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "22-R12")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "23-F121")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4D")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "22-R13")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "23-F72")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "23-F66")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4G")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4C")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank5.5C")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "23-F73")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank5.5G")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "23-F124")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4E")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank5.5H")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4F")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank5.5E")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank5.5D")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank5.5F")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4H ")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank6")
#
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank7")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank8")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank4")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank1")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank2")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank3")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "Blank5")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "22-R15")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "12578.blank4.4H")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "PCR_Pos1")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "PCR_Pos2")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "PCR_Neg1")
Metadata_reduced <- subset(Metadata_reduced, sample_name != "PCR_Neg2")
Metadata_reduced$sample_name<-gsub("-",".", as.character(Metadata_reduced$sample_name))
```
# Alpha Diversity {#alpha}

**Alpha diversity** measures the diversity within a single sample, reflecting the taxonomic richness and distribution of those microbial taxa groups within a sample.

**Note** that we are using the rarefied data for alpha diversity comparison. Rarefaction adjusts for varying library sizes, therefore adjusting subsamples to equal size to aid in alpha diversity comparison.

Prepare data for alpha diversity analyses
```{r alpha}

#Make richness table with diversity indices
alpha_full <- data.frame(estimate_richness(phyloseq_16S_rarifed, split = TRUE,
measures = NULL))
#
#alpha_full to alpha
alpha<-alpha_full
#
#create column in alpha to join by sample name
alpha<- tibble::rownames_to_column(alpha, "sample_name")
#
#Remove x from names
alpha$sample_name<-gsub("X","", as.character(alpha$sample_name))
row.names(alpha) <- alpha$sample_name
#
#join metadata to species richness data 

meta<-data.frame(sample_data(phyloseq_16S_rarifed))
#
#create a file to run ANOVA and run plots from
alpha_meta<- left_join(Metadata_reduced,alpha)
#alpha_meta<- merge(alpha,meta)
```
```{r alpha_table}

knitr::kable(alpha_meta, caption = "Table 3. Alpha Diversity Indices") %>%
    kable_styling(c("striped", "scale_down", "condensed")) %>% scroll_box(width = "700px", height = "800px")
deer<-subset(alpha_meta, Species !="Moose")
moose<-subset(alpha_meta, Species =="Moose")
#write.csv(deer, file = "DeerAlpha.csv")
#write.csv(moose, file = "MooseAlpha.csv")
```

## Alpha diversity back in the separate project folders

# Beta Diversity {#beta}

**Beta diversity** compares microbial diversity between samples.

Here, we will continue to used our rarefied data set. Prepare data for beta diversity analyses. We will look at data with rare taxa conserved (unfiltered), rare taxa removed (filtered), and log transformed.

```{r subset_for_beta}
moose_phylo<-subset_samples(phyloseq_16S_rarifed, Species=="Moose")
deer_phylo<-subset_samples(phyloseq_16S_rarifed, Species!="Moose")
INphylo<-subset_samples(moose_phylo, Location =="Inland NW") # separating out the two species up here because i didn't know where else to do it
AKphylo<-subset_samples(moose_phylo, Location =="Alaska")
MDphylo<-subset_samples(deer_phylo, Species =="Mule Deer") # separating out the two species up here because i didn't know where else to do it
WTDphylo<-subset_samples(deer_phylo, Species !="Mule Deer")
#moose_phylo<-data.frame(otu_table(moose_phylo))
```
**Prepare unfiltered data:**
```{r beta_rarefied_data_unfiltered}
#Transform to relative abundance
phyloseq_16S_tr_unfiltered <- transform_sample_counts(moose_phylo, 
function(x) x / sum(x))
#
#check to see that the sum for each column (each sample) is 1
sort(colSums(otu_table(phyloseq_16S_tr_unfiltered))) 
```

**Prepare filtered data:**
```{r beta_rarefied_data_filtered, echo=T, results='hide'}
#This next line subsets any taxa which have a mean abundance of less than 1E-5. 
phyloseq_16S_tr_filtered <- filter_taxa(phyloseq_16S_tr_unfiltered, function(x) mean(x) > 1e-5, TRUE)
```

**Prepare log-transformed data:**
```{r beta_rarefied_data_log, echo=T, results='hide'}
#Transform log, log1p(x)= log(1 +x) so as to not disrupt zero counts
phyloseq_16S_log<-transform_sample_counts(phyloseq_16S_tr_unfiltered, function(x) log1p(x))
```
## Ordinations

Ordination is a term for summarizing community data with the aim of transforming multivariate data in order to present it in fewer dimensions. There are many methods or series of methods of ordination.

The following code applies two ordination methods:

* **Principal Coordinates Analysis (PCoA) or Multidimensional scaling, MDS** - PCoA is a visualization method designed to study the similarity or difference of data based on distances other than Euclidean distances (as PCA uses). Instead, PCoA finds the potential principal components of the overall distance through dimensionality reduction. Therefore, PCoA can be used to study the similarity or difference of sample composition and observe the differences between groups.
** maximizes linear correlation between samples

* **Non-metric Multidimensional Scaling (NMDS)** - Multidimensional scaling is a means of visualizing the level of similarity of individual samples or sites of a multivariate data set (3+ variables) in a distance matrix. In comparison to metric MDS, non-metric MDS finds both a non-parametric monotonic (nonchanging; as one variable increases or decreases, the other variable does the same) relationship between the dissimilarities in the matrix and the Euclidean distances between

** maximizes rank-order (non-metric) correlation between samples

** rank-order approach well-suited for certain types of data, such as counts of abundance

with the following metrics:

* **Bray-Curtis** - Bray-Curtis is a statistical metric used in ecology to quantify the dissimilarity between sites based on counts per site. This allows us to examine the abundances of microbes shared common between sites, and the number of microbes in each. This metric ranges from 0 to 1. If both samples have the same number of microbes at the same abundance, the "dissimilarity" will equal 0.

* **unweighted UniFrac** - The unweighted UniFrac distance considers only species presence and absence information, counting the fraction of branch length unique to either community. (Chen, J. et al)

* **weighted UniFrac** - The weighted UniFrac distance considers the abundance information and weighs the branch length with abundance difference. (Chen, J. et al)
```{r beta_ordinations_unfiltered, results='hide'}
#Ordinate phylogenetic sequence data by PCoA, with distance method Bray-Curtis
bray_PCoA_unfiltered  <- ordinate(phyloseq_16S_tr_unfiltered, "PCoA", "bray")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method Bray-Curtis
bray_nmds_unfiltered  <- ordinate(phyloseq_16S_tr_unfiltered, "NMDS", "bray")
#
#Ordinate phylogenetic sequence data by PCoA, with distance method unweighted UniFrac
uni_PCoA_unfiltered <- ordinate(phyloseq_16S_tr_unfiltered, "PCoA", "unifrac")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method unweighted UniFrac
uni_nmds_unfiltered <- ordinate(phyloseq_16S_tr_unfiltered, "NMDS", "unifrac")
#
#Ordinate phylogenetic sequence data by PCoA, with distance method weighted UniFrac
wuni_PCoA_unfiltered <- ordinate(phyloseq_16S_tr_unfiltered, "PCoA", "wunifrac")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method weighted UniFrac
wuni_nmds_unfiltered <- ordinate(phyloseq_16S_tr_unfiltered, "NMDS", "wunifrac")
```

```{r beta_diversity_ordinations_Filtered, results='hide'}
##Create diversity matrices, this time without rare taxa 
#
#Ordinate phylogenetic sequence data by PCoA, with distance method Bray-Curtis
bray_PCoA  <- ordinate(phyloseq_16S_tr_filtered, "PCoA", "bray")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method Bray-Curtis
bray_nmds  <- ordinate(phyloseq_16S_tr_filtered, "NMDS", "bray")
#
#Ordinate phylogenetic sequence data by PCoA, with distance method unweighted UniFrac
uni_PCoA <- ordinate(phyloseq_16S_tr_filtered, "PCoA", "unifrac")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method unweighted UniFrac
uni_nmds <- ordinate(phyloseq_16S_tr_filtered, "NMDS", "unifrac")
#
#Ordinate phylogenetic sequence data by PCoA, with distance method weighted UniFrac
wuni_PCoA <- ordinate(phyloseq_16S_tr_filtered, "PCoA", "wunifrac")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method weighted UniFrac
wuni_nmds <- ordinate(phyloseq_16S_tr_filtered, "NMDS", "wunifrac")
```

```{r beta_diversity_ordinations_log_tr, results='hide'}
##Create diversity matrices, this time without rare taxa 
#
#Create diversity matrix, prefiltered relative log transformed log1p(x) 
bray_PCoA_log  <- ordinate(phyloseq_16S_log, "PCoA", "bray")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method Bray-Curtis
bray_nmds_log  <- ordinate(phyloseq_16S_log, "NMDS", "bray")
#
#Ordinate phylogenetic sequence data by PCoA, with distance method unweighted UniFrac
uni_PCoA_log <- ordinate(phyloseq_16S_log, "PCoA", "unifrac")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method unweighted UniFrac
uni_nmds_log <- ordinate(phyloseq_16S_log, "NMDS", "unifrac")
#
#Ordinate phylogenetic sequence data by PCoA, with distance method weighted UniFrac
wuni_PCoA_log <- ordinate(phyloseq_16S_log, "PCoA", "wunifrac")
#
#Ordinate phylogenetic sequence data by NMDS, with distance method weighted UniFrac
wuni_nmds_log <- ordinate(phyloseq_16S_log, "NMDS", "wunifrac")
```


## Plots for Beta Diversity Ordinations

Create beta diversity plots utilizing the matrices established above.

### Unfiltered Data

#### Unfiltered Data: Study Site

Plot PCoA, Bray-Curtis:
```{r beta_site_bray_pcoa_plot_unfiltered}
beta_otu<-sort(colSums(otu_table(moose_phylo)))
factor(sample_data(phyloseq_16S_tr_unfiltered)$Sample_Type)
levels(sample_data(phyloseq_16S_tr_unfiltered)$Sample_Type) = ordered(c("Rumen", "Cecum", "Colon"))

Conc_bray_PCoA_plot2 <- plot_ordination(phyloseq_16S_tr_unfiltered, 
                                         bray_PCoA_unfiltered,
                                         color = "Sample_Type",
                                        shape = "Location") + theme_bw()+ theme(text=element_text(size=16))
  
Conc_bray_PCoA_plot2 + stat_ellipse(linewidth=.85)+
  geom_point(size=4)+
  scale_color_manual(values=c("lightskyblue3", "darkseagreen4","gray4"))
  #xlim(-0.5,0.5)+
  #ylim(-0.6s,0.6)

```


Plot PCoA, Unweighted UniFrac:
```{r beta_site_uni_pcoa_plot_unfiltered}
site_uni_PCoA_unfiltered <- plot_ordination(phyloseq_16S_tr_unfiltered, 
                                             uni_PCoA_unfiltered,
                                            color = "Sample_Type",
                                        shape = "Location") + theme_bw()+ theme(text=element_text(size=16))
site_uni_PCoA_unfiltered + stat_ellipse() + 
  geom_point(size=4)+
  scale_color_manual(values=c("lightskyblue3", "darkseagreen","gray41"))
```

Plot PCoA, Weighted UniFrac:
```{r beta_site_wuni_pcoa_plot_unfiltered}
site_wuni_PCoA_unfiltered <- plot_ordination(phyloseq_16S_tr_unfiltered, 
                                              wuni_PCoA_unfiltered,
                          color = "Trial", shape = "Species") + theme_bw()
site_wuni_PCoA_unfiltered + stat_ellipse() + 
  labs(title = "Weighted Unifrac Dissimilarity PCoA Study Site, Unfiltered") + 
  theme(text = element_text(size = 10))

```

Plot NMDS, Bray-Curtis:
```{r beta_site_bray_nmds_plot_unfiltered}
site_bray_NMDS_unfiltered <- plot_ordination(phyloseq_16S_tr_unfiltered, 
                                              bray_nmds_unfiltered,
                                                color = "Trial", shape = "Species") + theme_bw()

site_bray_NMDS_unfiltered + stat_ellipse() + 
  labs(title = "Bray-Curtis Dissimilarity NMDS Study Site, Unfiltered") + 
  theme(text = element_text(size = 10))
```

Plot NMDS, Unweighted UniFrac: 
```{r beta_site_uni_nmds_plot_unfiltered}
site_uni_nmds_unfiltered <- plot_ordination(phyloseq_16S_tr_unfiltered, 
                                         uni_nmds_unfiltered,
                                         type = "Location",
                                             color = "Sample_Type") + theme_bw()
  
site_uni_nmds_unfiltered + stat_ellipse() + 
  labs(title = "Unifrac Dissimilarity NMDS Study Site, Unfiltered") + 
  theme(text = element_text(size = 10))

```

Plot NMDS, Weighted Unifrac:
```{r site_wuni_nmds_plot_unfiltered}
site_wuni_PCoA_plot2 <- plot_ordination(phyloseq_16S_tr_unfiltered, 
                                         wuni_nmds_unfiltered,
                                       color = "Trial", shape = "Species") + theme_bw()
  
site_wuni_PCoA_plot2 + stat_ellipse() + 
  labs(title = "Weighted Unifrac Dissimilarity NMDS Study Site, Unfiltered") + 
  theme(text = element_text(size = 10))
```

# Taxonomy {#tax}

Prepare data for taxonomy analysis. Since taxonomy was matched with the Silva database, we can compare what microbes are in the fecal and cecal samples, and how taxa compares between the fecal and cecal pellets, and two locations.
# Permanova Tests
```{r perm}

#Setup tests with transformed, pre-filtered files
#Pull otu_table out of phyloseq object
table_gp <- data.frame(otu_table(phyloseq_16S_tr_filtered)) 

#Transpose data
flip_ps_table_gp<-t(table_gp)

#Pull the sample_data metadata out of phyloseq object 
meta_gp<- data.frame(sample_data(phyloseq_16S_tr_filtered))
#check
#flip_ps_table_gp [1:20, 1:10]
#check
#View(meta_gp) 
#Create data matrices
adonis_table<-data.matrix(flip_ps_table_gp)
adon<-merge(adonis_table, meta_gp)

#Set seed for study site
set.seed(223)
adon<-adonis2((adonis_table)~+meta_gp$Location, na.action=na.omit)
print(adon)

simple.results.anosim <- anosim(x = adonis_table, grouping = meta_gp$Sample_Type, permutations = 999)
simple.results.anosim
```
```{r new pcoa}
# Load required libraries
library(vegan)   # for vegdist
library(ape)     # for pcoa
library(ggplot2)
library(dplyr)

# Assume you already have:
# - otu_table: samples x taxa matrix (rows = samples, columns = taxa)
# - metadata: dataframe with SampleID and Time (numeric or factor with ordering)

# Step 1: Compute Bray-Curtis distance matrix
dist_matrix <- vegdist(beta_otu, method = "bray") #og code vegdist(otu_table, method = "bray") used the sort function from above (where I separated the two species) to make the otu table 

# Step 2: Run PCoA
pcoa_result <- pcoa(dist_matrix)

# Step 3: Extract first 2 PCoA axes
pcoa_scores <- as.data.frame(pcoa_result$vectors[, 1:2])
pcoa_scores$sample_name <- rownames(pcoa_scores)

# Step 4: Merge with metadata
pcoa_scores <- left_join(pcoa_scores, meta_gp, by = "sample_name")

# Step 5: Compute centroids for each time point
centroids <- pcoa_scores %>%
  group_by(Location) %>%
  group_by(Sample_Type)%>%
  summarise(
    PCoA1 = mean(Axis.1),
    PCoA2 = mean(Axis.2),
    .groups = "drop"
  )

# Step 6: Plot with trajectories

centroids <- pcoa_scores %>%
  group_by(Location, Sample_Type) %>%
  summarize(PCoA1 = mean(Axis.1),PCoA2 = mean(Axis.2))

ggplot(centroids, aes(x = PCoA1, y = PCoA2)) +
  geom_point(aes(shape = Location, color = Sample_Type), size = 5) +
  #geom_text(vjust = -1.5, size = 5) +
  geom_path(aes(group = Location), arrow = arrow(length = unit(0.3, "cm")), linetype = "solid") +
  theme_minimal()+
  scale_color_manual(values = c("lightskyblue3", "darkseagreen4","gray4"))+
 theme_bw()+ theme(text=element_text(size=16))
```
#Pairwise comparisions

```{r, include=FALSE}

BC.dist=vegdist(adonis_table, distance="bray")
disp.type = betadisper(BC.dist, meta_gp$Sample_Type)
permutest(disp.type, pairwise=TRUE, permutations=1000)
location = betadisper(BC.dist, meta_gp$Location)
permutest(location, pairwise=TRUE, permutations=1000)
```


## Phyla

The following code was adapted from Courtney Buchanan's code:

```{r phyla_relative_abundance, warning=FALSE}
# Group all ASVs by phylum
phyloseq_16S_phylum <- tax_glom(moose_phylo, taxrank = "Phylum")
## Test % in each level top 10, 15, 20
#
# Identify the top 10 phylum 
phylum_10 <- names(sort(taxa_sums(phyloseq_16S_phylum), TRUE)[1:10])
#
# Keep data for only the top 10 phyla
phyloseq_16S_phylum_prune10 <- prune_taxa(phylum_10, phyloseq_16S_phylum)
#
# Identify the top 15 phyla 
phylum_15 <- names(sort(taxa_sums(phyloseq_16S_phylum), TRUE)[1:15])
#
# Keep data for only the top 15 phyla
phyloseq_16S_phylum_prune15 <- prune_taxa(phylum_15, phyloseq_16S_phylum)
#
# Identify the top 20 phyla 
phylum_20 <- names(sort(taxa_sums(phyloseq_16S_phylum), TRUE)[1:20])
#
# Keep data for only the top 20 phyla
phyloseq_16S_phylum_prune20 <- prune_taxa(phylum_20, phyloseq_16S_phylum)
#
# Check percentage of reads that fall within top 10 phyla, 15, 20 
sum(sample_sums(phyloseq_16S_phylum_prune10)) / sum(sample_sums(phyloseq_16S_phylum))
sum(sample_sums(phyloseq_16S_phylum_prune15)) / sum(sample_sums(phyloseq_16S_phylum))
sum(sample_sums(phyloseq_16S_phylum_prune20)) / sum(sample_sums(phyloseq_16S_phylum))
#
# top 10 captures most, so we will use this below
# Make table
phylum_all_table <- cbind(tax_table(phyloseq_16S_phylum))
#
#trying to re-name some phyla to other....
phylum_new_table <- phylum_all_table
phylum_new_table<- tibble::rownames_to_column((as.data.frame(phylum_new_table)), "ASV") #add ASV as column
#
phylum_new_table$Phylum1 <- phylum_new_table$Phylum #makes new column for renaming
#
# rename top 10
phylum_new_table$Phylum1[phylum_new_table$ASV == "cee47dbd6c7810e033fda078a78d377d"] <- "top_10" #1
phylum_new_table$Phylum1[phylum_new_table$ASV == "ac164649e33237f799c1ffee4f344d5d"] <- "top_10" #2
phylum_new_table$Phylum1[phylum_new_table$ASV == "beaaaff5eb38cae1fedca5779ebdca3f"] <- "top_10" #3
phylum_new_table$Phylum1[phylum_new_table$ASV == "ecaa0f1d2a71e76b2c58c4787ca10b41"] <- "top_10" #4
phylum_new_table$Phylum1[phylum_new_table$ASV == "e2a9ad58bbca06a4f08085840783fbcd"] <- "top_10" #5
phylum_new_table$Phylum1[phylum_new_table$ASV == "ba769481e813de83cbe84678905718de"] <- "top_10" #6
phylum_new_table$Phylum1[phylum_new_table$ASV == "dd864abcc827d7c8d52f7156f65707ad"] <- "top_10" #7
phylum_new_table$Phylum1[phylum_new_table$ASV == "d0e8b75ff5314350e4a97dd3db138f1b"] <- "top_10" #8
phylum_new_table$Phylum1[phylum_new_table$ASV == "429ad9f2ff02c3a27527d7f2d0cbc9fe"] <- "top_10" #9
phylum_new_table$Phylum1[phylum_new_table$ASV == "f33c1ca6142229599fbd1700ac708c25"] <- "top_10" #10
#
phylum_new_table$Phylum[phylum_new_table$Phylum1 !="top_10"] <- "Additional Phyla"
#
#make ASV row name again
phylum_new_table1 <- tibble::column_to_rownames((as.data.frame(phylum_new_table)), "ASV") 
#
#need to remove Phylum1 column
phylum_relabeled <- phylum_new_table1
phylum_relabeled = subset(phylum_new_table1, select=-c(Phylum1))
#
phylum_relabeled1 <- as.matrix(phylum_relabeled)#create a matrix
str(phylum_relabeled1)
#try to merge into the phylum classified phyloseq object 
phyloseq_16S_phylum_sort<-phyloseq_16S_phylum #create new object
#
#need to be able to upload as new taxonomy table 
tax_table(phyloseq_16S_phylum_sort) <- phylum_relabeled1
```
### Phyla: Study Site


## Class

```{r class_relative_abundance, warning=FALSE}
# Group all ASVs by class
phyloseq_16S_class <- tax_glom(moose_phylo, taxrank = "Class")
#
# Identify the top 10 class 
class_10 <- names(sort(taxa_sums(phyloseq_16S_class), TRUE)[1:10])
#
# Keep data for only the top 10 class
phyloseq_16S_class_prune10 <- prune_taxa(class_10, phyloseq_16S_class)
#
# Identify the top 15 class
class_15 <- names(sort(taxa_sums(phyloseq_16S_class), TRUE)[1:15])
#
# Keep data for only the top 15 class
phyloseq_16S_class_prune15 <- prune_taxa(class_15, phyloseq_16S_class)
#
# Identify the top 20 class
class_20 <- names(sort(taxa_sums(phyloseq_16S_class), TRUE)[1:20])
#
# Keep data for only the top 20 class
phyloseq_16S_class_prune20 <- prune_taxa(class_20, phyloseq_16S_class)
#
# Check percentage of reads that fall within top 10, 15, 20 class
sum(sample_sums(phyloseq_16S_class_prune10)) / sum(sample_sums(phyloseq_16S_class))
sum(sample_sums(phyloseq_16S_class_prune15)) / sum(sample_sums(phyloseq_16S_class))
sum(sample_sums(phyloseq_16S_class_prune20)) / sum(sample_sums(phyloseq_16S_class))
#
# Most diversity captured in top 10, so using this moving forward. 
#
# Make table
class_all_table <- cbind(tax_table(phyloseq_16S_class))
#
#renaming somw classes to "other"
class_new_table <- class_all_table
class_new_table <- tibble::rownames_to_column((as.data.frame(class_new_table)), "ASV") #add ASV as column
#
class_new_table$Class1 <- class_new_table$Class#makes new column for renaming
#
# rename top 10
class_new_table$Class1[class_new_table$ASV == "cee47dbd6c7810e033fda078a78d377d"] <- "top_10" #1
class_new_table$Class1[class_new_table$ASV == "ac164649e33237f799c1ffee4f344d5d"] <- "top_10" #2
class_new_table$Class1[class_new_table$ASV == "beaaaff5eb38cae1fedca5779ebdca3f"] <- "top_10" #3
class_new_table$Class1[class_new_table$ASV == "2642a3f84f710a36e4bc83095337f698"] <- "top_10" #4
class_new_table$Class1[class_new_table$ASV == "ecaa0f1d2a71e76b2c58c4787ca10b41"] <- "top_10" #5
class_new_table$Class1[class_new_table$ASV == "e2a9ad58bbca06a4f08085840783fbcd"] <- "top_10" #6
class_new_table$Class1[class_new_table$ASV == "384962c69cdd7b9e3bd01f030143b0b6"] <- "top_10" #7
class_new_table$Class1[class_new_table$ASV == "dd864abcc827d7c8d52f7156f65707ad"] <- "top_10" #8
class_new_table$Class1[class_new_table$ASV == "d0e8b75ff5314350e4a97dd3db138f1b"] <- "top_10" #9
class_new_table$Class1[class_new_table$ASV == "ba769481e813de83cbe84678905718de"] <- "top_10" #10
#
class_new_table$Class[class_new_table$Class1 !="top_10"] <- "Additional Classes"
#
#make ASV row name again
class_new_table1<- tibble::column_to_rownames((as.data.frame(class_new_table)), "ASV") 
#
#need to remove Class1 column
class_relabeled<-class_new_table1
class_relabeled= subset(class_new_table1, select=-c(Class1))
class_relabeled1 <- as.matrix(class_relabeled)#create a matrix and maybe that will work?
#try to merge into the phylum classified phyloseq object 
phyloseq_16S_class_sort<-phyloseq_16S_class #create new object
#new taxonomy table
tax_table(phyloseq_16S_class_sort) <- class_relabeled1
```


Here, we will plot relative abundance for order. For visualization purposes, we will look at the top 10 orders as most microbes fell into these categories. 
```{r order_relative_abundance, warning=FALSE}
# Group all ASVs by order
phyloseq_16S_order <- tax_glom(moose_phylo, taxrank = "Order")
# 
#Run the following line to see how many taxa exist. For example, for order, there are 132.
phyloseq_16S_order
#
## Identify all orders. Be sure to adjust the range. For order, since there are 132 orders, adjust the range to capture all 132 orders.
order_all <- names(sort(taxa_sums(phyloseq_16S_order), TRUE)[1:132])
#
# Identify the top 10 order 
order_10 <- names(sort(taxa_sums(phyloseq_16S_order), TRUE)[1:10])
#
# Keep data for only the top 10 order in this object
phyloseq_16S_order_prune10 <- prune_taxa(order_10, phyloseq_16S_order)
#
# Check what percentage of reads fall in the top 10 orders. Note that we can do this with the top 15, 20, etc. as well.
sum(sample_sums(phyloseq_16S_order_prune10)) / sum(sample_sums(phyloseq_16S_order))
#
# Make table
order_all_table <- cbind(tax_table(phyloseq_16S_order))
#
#Make Table
order_new_table <- order_all_table
order_new_table<- tibble::rownames_to_column((as.data.frame(order_new_table)), "ASV") #add ASV as column
#
order_new_table$Order1<-order_new_table$Order #makes new column for renaming
#
# rename to top 10
order_new_table$Order1[order_new_table$ASV == "ac164649e33237f799c1ffee4f344d5d"] <- "top_10" #1
order_new_table$Order1[order_new_table$ASV == "cee47dbd6c7810e033fda078a78d377d"] <- "top_10" #2
order_new_table$Order1[order_new_table$ASV == "beaaaff5eb38cae1fedca5779ebdca3f"] <- "top_10" #3
order_new_table$Order1[order_new_table$ASV == "80ffd60a547eca9b1cb49df58b3b209f"] <- "top_10" #4
order_new_table$Order1[order_new_table$ASV == "2bdd097da214d4453890af6d34b2452f"] <- "top_10" #5
order_new_table$Order1[order_new_table$ASV == "04235b7635e60adc8cae82a3f3b2b8d3"] <- "top_10" #6
order_new_table$Order1[order_new_table$ASV == "8b77f352fd8d25b662238c15c05fa6b2"] <- "top_10" #7
order_new_table$Order1[order_new_table$ASV == "ecaa0f1d2a71e76b2c58c4787ca10b41"] <- "top_10" #8
order_new_table$Order1[order_new_table$ASV == "2b1f51d9406df70bb9a4a499bace04c9"] <- "top_10" #9
order_new_table$Order1[order_new_table$ASV == "38370559a53781fc8c5f6aaa5323d956"] <- "top_10" #10
order_new_table$Order[order_new_table$Order1 !="top_10"] <- "Additional Orders"
#
#make ASV row name again
order_new_table1<- tibble::column_to_rownames((as.data.frame(order_new_table)), "ASV") 
#
#remove Order1 column
order_relabeled<-order_new_table1
order_relabeled= subset(order_new_table1, select=-c(Order1))
#
order_relabeled1 <- as.matrix(order_relabeled)
#
#try to merge into the Class classified phyloseq object 
phyloseq_16S_order_sort<-phyloseq_16S_order#create new object so I dont ruin the 
#
#need to be able to upload as new taxonomy table 
tax_table(phyloseq_16S_order_sort) <- order_relabeled1
```


```{r, include=FALSE}

# top 10 captures most, so we will use this below
phyloseq_16S_fam <- tax_glom(moose_phylo, taxrank = "Family")
phyloseq_In <- subset_samples(phyloseq_16S_fam, Location == "Inland NW")
phyloseq_AK <- subset_samples(phyloseq_16S_fam, Location == "Alaska")

## Identify top 10 Phyla
phyloseq_In_20 <- names(sort(taxa_sums(phyloseq_In), TRUE)[1:20])
phyloseq_In_20
phyloseq_AK_20 <- names(sort(taxa_sums(phyloseq_AK), TRUE)[1:20])
phyloseq_AK_20
## Keep data for only top 10 families
phyloseq_IN_prune<- prune_taxa(phyloseq_In_20, phyloseq_In)
phyloseq_AK_prune <- prune_taxa(phyloseq_AK_20, phyloseq_AK)
# Make table
all_table1 <- cbind(tax_table(phyloseq_IN_prune))
all_table2 <- cbind(tax_table(phyloseq_AK_prune))
#trying to re-name some phyla to other....
fam_all_table1 <- all_table1
fam_all_table1<- tibble::rownames_to_column((as.data.frame(fam_all_table1)), "ASV") #add ASV as column
View(fam_all_table1)#check

fam_all_table1$Family1 <- fam_all_table1$Family #makes new column for renaming
fam_all_table1$Family1[fam_all_table1$ASV == "2bdd097da214d4453890af6d34b2452f"] <- "top_20" #1
fam_all_table1$Family1[fam_all_table1$ASV == "3091aa0dbc7fbe655f65b51f61c78d54"] <- "top_20" #2
fam_all_table1$Family1[fam_all_table1$ASV == "4d20c23b6a8ea0b70c925e43bf5901d8"] <- "top_20" #3
fam_all_table1$Family1[fam_all_table1$ASV == "3102de2a43e0034c0d5ea30f4f178d6f"] <- "top_20" #4
fam_all_table1$Family1[fam_all_table1$ASV == "8b77f352fd8d25b662238c15c05fa6b2"] <- "top_20" #5
fam_all_table1$Family1[fam_all_table1$ASV == "8a4e873ffb564107c3e27056cafbd42f"] <- "top_20" #6
fam_all_table1$Family1[fam_all_table1$ASV == "cee47dbd6c7810e033fda078a78d377d"] <- "top_20" #7
fam_all_table1$Family1[fam_all_table1$ASV == "fe65890615d238fd02383c9ca3792318"] <- "top_20" #8
fam_all_table1$Family1[fam_all_table1$ASV == "2642a3f84f710a36e4bc83095337f698"] <- "top_20" #9
fam_all_table1$Family1[fam_all_table1$ASV == "a4ad22d7dad0d45debdce30a48b3b02a"] <- "top_20" #10
fam_all_table1$Family1[fam_all_table1$ASV == "f645be84f45c4030599aa3c00ad15a19"] <- "top_20" #11
fam_all_table1$Family1[fam_all_table1$ASV == "04235b7635e60adc8cae82a3f3b2b8d3"] <- "top_20" #12
fam_all_table1$Family1[fam_all_table1$ASV == "d3313b86bca203789bf987d5aaa0e78b"] <- "top_20" #13
fam_all_table1$Family1[fam_all_table1$ASV == "80ffd60a547eca9b1cb49df58b3b209f"] <- "top_20" #14
fam_all_table1$Family1[fam_all_table1$ASV == "ecaa0f1d2a71e76b2c58c4787ca10b41"] <- "top_20" #15
fam_all_table1$Family1[fam_all_table1$ASV == "41f917636a6ff526166a8163bb2b1e96"] <- "top_20" #16
fam_all_table1$Family1[fam_all_table1$ASV == "e2a9ad58bbca06a4f08085840783fbcd"] <- "top_20" #17
fam_all_table1$Family1[fam_all_table1$ASV == "bedab54808bec7ba26224776addd4693"] <- "top_20" #18
fam_all_table1$Family1[fam_all_table1$ASV == "cf770b15dd490c11199c571f5376c398"] <- "top_20" #19
fam_all_table1$Family1[fam_all_table1$ASV == "ac164649e33237f799c1ffee4f344d5d"] <- "top_20" #20
View(fam_all_table1) # check and it worked
fam_all_table1$Family[fam_all_table1$Family1 !="top_20"] <- "Additional Families"
View(fam_all_table1) #check
#make ASV row name again
fam_all_table1<- tibble::column_to_rownames((as.data.frame(fam_all_table1)), "ASV") 
#need to remove Order1 coulumn
fam_relabeled1<-fam_all_table1
fam_relabeled1= subset(fam_all_table1, select=-c(Family1))
View(fam_relabeled1) #check-yes it worked
str(fam_relabeled1)#check structure
fam_relabeled1 <- as.matrix(fam_relabeled1)#create a matrix and maybe that will work?
str(fam_relabeled1)#check structure now
#try to merge into the phylum classified phyloseq object 
phyloseq_16S_fam_sort1<-phyloseq_In#create new object so I dont ruin the old one...
phyloseq_16S_fam_sort1
#need to be able to upload as new taxonomy table 
tax_table(phyloseq_16S_fam_sort1) <- fam_relabeled1 #try matrix one... it seemed to work 
#Once I get formatting should work here on.... 
# Transform abundances to percentages
fam_all_merge1 <- merge_samples(phyloseq_16S_fam_sort1, "Sample_Type")
fam_all_percent1 <- transform_sample_counts(fam_all_merge1, function(x) 100 * x/sum(x))
fam_all_percent1
fam_abundance_sort1 <- plot_bar(fam_all_percent1, fill = "Family") + xlab("Gastrointestinal Organ") + ylab("Relative Abundance (%)") +  
    scale_x_discrete(labels = c("Rumen", "Cecum", "Colon"))+
     scale_fill_manual(values=natparks.pals("RockyMtn", 20)) +
    ylab("Relative Abundance (%)") + 
    ggtitle("Inland NW")+
    theme_bw()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
fam_abundance_sort1 + theme(text = element_text(size = 12))
#########################################################################################################################################################
fam_all_table2 <- all_table2
fam_all_table2<- tibble::rownames_to_column((as.data.frame(fam_all_table2)), "ASV") #add ASV as column
View(fam_all_table2)#check

fam_all_table2$Family1 <- fam_all_table2$Family #makes new column for renaming
fam_all_table2$Family1[fam_all_table2$ASV == "2bdd097da214d4453890af6d34b2452f"] <- "top_20" #1
fam_all_table2$Family1[fam_all_table2$ASV == "3091aa0dbc7fbe655f65b51f61c78d54"] <- "top_20" #2
fam_all_table2$Family1[fam_all_table2$ASV == "4d20c23b6a8ea0b70c925e43bf5901d8"] <- "top_20" #3
fam_all_table2$Family1[fam_all_table2$ASV == "3102de2a43e0034c0d5ea30f4f178d6f"] <- "top_20" #4
fam_all_table2$Family1[fam_all_table2$ASV == "8a4e873ffb564107c3e27056cafbd42f"] <- "top_20" #5
fam_all_table2$Family1[fam_all_table2$ASV == "e14f35e77d491dd6e8a7619f5f3b8c45"] <- "top_20" #6
fam_all_table2$Family1[fam_all_table2$ASV == "cee47dbd6c7810e033fda078a78d377d"] <- "top_20" #7
fam_all_table2$Family1[fam_all_table2$ASV == "fe65890615d238fd02383c9ca3792318"] <- "top_20" #8
fam_all_table2$Family1[fam_all_table2$ASV == "2b1f51d9406df70bb9a4a499bace04c9"] <- "top_20" #9
fam_all_table2$Family1[fam_all_table2$ASV == "dd864abcc827d7c8d52f7156f65707ad"] <- "top_20" #10
fam_all_table2$Family1[fam_all_table2$ASV == "04235b7635e60adc8cae82a3f3b2b8d3"] <- "top_20" #11
fam_all_table2$Family1[fam_all_table2$ASV == "d3313b86bca203789bf987d5aaa0e78b"] <- "top_20" #12
fam_all_table2$Family1[fam_all_table2$ASV == "38370559a53781fc8c5f6aaa5323d956"] <- "top_20" #13
fam_all_table2$Family1[fam_all_table2$ASV == "74f360c17fd5551faba788990ba35e8d"] <- "top_20" #14
fam_all_table2$Family1[fam_all_table2$ASV == "80ffd60a547eca9b1cb49df58b3b209f"] <- "top_20" #15
fam_all_table2$Family1[fam_all_table2$ASV == "c1e1874e83a9e6bd9509862ec9a25805"] <- "top_20" #16
fam_all_table2$Family1[fam_all_table2$ASV == "c6f1e03b8fc0bee210607216c56699a9"] <- "top_20" #17
fam_all_table2$Family1[fam_all_table2$ASV == "beaaaff5eb38cae1fedca5779ebdca3f"] <- "top_20" #18
fam_all_table2$Family1[fam_all_table2$ASV == "cf770b15dd490c11199c571f5376c398"] <- "top_20" #19
fam_all_table2$Family1[fam_all_table2$ASV == "ac164649e33237f799c1ffee4f344d5d"] <- "top_20" #20
View(fam_all_table2) # check and it worked
fam_all_table2$Family[fam_all_table2$Family1 !="top_20"] <- "Additional Families"
View(fam_all_table2) #check
#make ASV row name again
fam_new_table2<- tibble::column_to_rownames((as.data.frame(fam_all_table2)), "ASV") 
#need to remove Order1 coulumn
fam_relabeled2<-fam_new_table2
fam_relabeled2= subset(fam_new_table2, select=-c(Family1))
View(fam_relabeled2) #check-yes it worked
str(fam_relabeled2)#check structure
fam_relabeled2 <- as.matrix(fam_relabeled2)#create a matrix and maybe that will work?
str(fam_relabeled2)#check structure now
#try to merge into the phylum classified phyloseq object 
phyloseq_16S_fam_sort2<-phyloseq_AK#create new object so I dont ruin the old one...
phyloseq_16S_fam_sort2
#need to be able to upload as new taxonomy table 
tax_table(phyloseq_16S_fam_sort2) <- fam_relabeled2 #try matrix one... it seemed to work 
#Once I get formatting should work here on.... 
# Transform abundances to percentages
fam_all_merge2 <- merge_samples(phyloseq_16S_fam_sort2, "Sample_Type")
fam_all_percent2 <- transform_sample_counts(fam_all_merge2, function(x) 100 * x/sum(x))
fam_all_percent2

fam_abundance_sort2 <- plot_bar(fam_all_percent2, fill = "Family") + xlab("Gastrointestinal Organ") + ylab("Relative Abundance (%)") +  
     scale_x_discrete(labels = c("Rumen", "Cecum", "Colon"))+
        scale_fill_manual(values=natparks.pals("RockyMtn", 20)) +
    ylab("Relative Abundance (%)") + 
    ggtitle("Alaska")+
    theme_bw()+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
fam_abundance_sort2 + theme(text = element_text(size = 12))


fam_abundance_sort1
fam_abundance_sort2
```

```{r, include=FALSE}
library(tidytext)
library("openxlsx") 

percentages_In<-psmelt(fam_all_percent1)
percentages_AK<-psmelt(fam_all_percent2)
attempt<- data.frame(sample_data(percentages_In))
Top10both2=createWorkbook() # create a new workbook for the final dataframe
addWorksheet(Top10both2,sheetName="IN")
writeData(Top10both2,percentages_In,sheet="IN",startRow=1,startCol=1,colNames=TRUE)
addWorksheet(Top10both2,sheetName="AK")
writeData(Top10both2,percentages_AK,sheet="AK",startRow=1,startCol=1,colNames=TRUE)
Fname=paste0(c("Top10both2.xlsx"),collapse="")
saveWorkbook(Top10both2,Fname, overwrite=FALSE)
```
```{r abundances plots}
IN_Plot  <- ggplot(data= percentages_In, aes(x=Sample,y=Abundance, fill=reorder(Family, Abundance)))+ 
  scale_x_discrete(labels = c("Rumen", "Cecum", "Colon"))+
   geom_bar( stat="identity", position="stack", colour = "black")+
  scale_fill_manual("Families Identified", values=natparks.pals("RockyMtn", 20)) +
 ylab("Relative Abundance (%)") + 
    xlab("Gastrointestinal Organ") + 
    ggtitle("Inland NW")+
    theme_bw()+
   theme(text=element_text(size=16), axis.text.x = element_text(angle = 45, hjust = 1))

AK_Plot  <- ggplot(data= percentages_AK, aes(x=Sample,y=Abundance, fill=reorder(Family, Abundance)))+ 
  scale_x_discrete(labels = c("Rumen", "Cecum", "Colon"))+
   geom_bar( stat="identity", position="stack", colour = "black")+
  scale_fill_manual("Families Identified", values=natparks.pals("RockyMtn", 20)) +
   ylab("Relative Abundance (%)") + 
    xlab("Gastrointestinal Organ") + 
    ggtitle("Alaska")+
    theme_bw()+
   theme(text=element_text(size=16), axis.text.x = element_text(angle = 45, hjust = 1))
 
IN_Plot
AK_Plot
```
# References

<div id="refs"></div>

---
nocite: '@*'
...

# Appendix 1

Here you will find citations of all R packages used to generate this report. This appendix was generated using the package `knitcitations` and command `read.bibtex()`.

```{r r_packages_bibliography, results = "asis", echo = FALSE, warning = FALSE, message = FALSE}
#
# Load associated R package
library("knitcitations")
#
# Prepare to print all citations in "packages.bib" file by clearing all bibliographies that could be in the cache
cleanbib()
#
# Set pandoc as the default output option for bib
options(citation_format = "pandoc")
#
# Read and print bibliography from "packages.bib" file generated in the packages code chunk
read.bibtex(file = "packages.bib")
```

# Appendix 2

Here you will find information about the version of R and operating system (OS) used, as well as the versions of loaded R packages. This appendix was generated using `sessionInfo()`.

```{r package_versions, echo=FALSE}
# Current R session and package versions
sessionInfo()
```
